A boundary-integral framework to simulate \\ viscous erosion of a porous medium

### 2018-04-09 docs/methods/erosion_methods.tex
NM

\title{A boundary-integral framework to simulate \\ the viscous erosion of a porous medium}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction\label{s:intro}}
Flow-induced erosion deteriorates and reshapes solid material over a range of scales found in nature, from large geological formations~\cite{han1969, Rothman2012, coh-dev-sey-yi-szy-rot2015}, down to human-scale features and patterns~\cite{daerr2003erosion}. Though less visible, these same forces work at the very smallest scales, slowly deteriorating the individual constituents of porous media (e.g.~solid, sand, or clay) or biological structures like plaque and biofilms~\cite{pic-van-hei2000, sha2002, gro-gij-van-fer-hat-van-yua-wen2007}. Motivated by such examples, we develop in this paper numerical methods to simulate fluid-mechanical erosion of multiple solid bodies in a Stokes flow---the most relevant regime for groundwater and the aforementioned bio-fluid applications.

Recent work in the high-Reynolds-number regime has highlighted the importance of a shape-flow feedback that occurs during erosion~\cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013, hewett2017evolution, moore2017riemann} and related processes such as dissolution and melting~\cite{Huang2015, kondratiuk2015steadily, rycroft2016asymmetric, cohen2016erosion, hewett2017pear, claudin2017dissolution}.  In these cases, the flow alters the morphology of the immersed structures, which in turn modifies the surrounding flow, and so on. Such processes of `sculpting with flow' \cite{ristroph2018sculpting} have been examined in the context of a single body \cite{ris-moo-chi-she-zha2012} and, more recently, multiple bodies whose shape evolution is coupled through the flow \cite{hewett2017evolution, moore2017riemann}.
From this viewpoint, the example of an eroding porous medium presents new opportunities and challenges. First, since the medium is composed of many individual constituents, multi-body interactions must be considered. Second, the long-range effects inherent to Stokes flow implies that the shape-evolution of bodies are even more strongly coupled than in any other regime.

@@

In closely related work, Mitchell \& Spagnolie used a traction integral equation formulation to simulate viscous erosion of a single body in three dimensions \cite{mit-spa2016}. Their formulation was flexible enough to simulate a {\em mobile} body that erodes in various background flows (uniform flow, shear flow, etc.). An advantage of the traction formulation is that is directly provides the shear stress that is needed to evolve surfaces. 
%
Since our aim is to simulate porous-medium erosion, we will consider {\em many} bodies and, further,  we choose to fix these bodies to their initial locations. Thus, instead of a single, mobile body, we will consider multiple, immobile bodies.
Rather than the traction-formulation, we will employ a standard double-layer formulation which will easily allow us to compute bulk quantities such as pressure and vorticity as a post-processing step.

%That work employed a traction integral equation to solve for the surrounding Stokes flow. The advantage to this formulation is that it immediately provides the shear stress required to evolve boundaries. However, the standard double-layer potential used here will allow us to compute certain quantities in the fluid bulk, such as vorticity and pressure, that will facilitate our understanding of the erosion process. Furthermore, our primary aim is to simulate the erosion of multiple bodies held fixed within a porous network, whereas Mitchell \& Spagnolie simulated a single eroding body that is mobile.

>>>>

\title{A boundary-integral framework to simulate \\ viscous erosion of a porous medium}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction\label{s:intro}}
Flow-induced erosion deteriorates and reshapes solid material over a range of scales found in nature, from large geological formations~\cite{han1969, Rothman2012, coh-dev-sey-yi-szy-rot2015}, down to human-scale features and patterns~\cite{daerr2003erosion}. Though less visible, these same forces are working at the very smallest scales, slowly deteriorating the individual constituents of porous media (e.g.~solid, sand, or clay) or biological structures like plaque and biofilms~\cite{pic-van-hei2000, sha2002, gro-gij-van-fer-hat-van-yua-wen2007}. Motivated by such examples, this paper develops numerical methods to simulate fluid-mechanical erosion of multiple solid bodies in a Stokes flow---the most relevant regime for groundwater and the aforementioned bio-fluid applications.

Recent work in the high-Reynolds-number regime has highlighted the importance of a shape-flow feedback that occurs during erosion~\cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013, hewett2017evolution, moore2017riemann} and the related processes of dissolution and melting~\cite{Huang2015, kondratiuk2015steadily, rycroft2016asymmetric, cohen2016erosion, hewett2017pear, claudin2017dissolution}.  In these cases, the flow alters the morphology of the immersed structures, which in turn modifies the surrounding flow, and so on. Such processes of `sculpting with flow' \cite{ristroph2018sculpting} have been examined in the context of a single body \cite{ris-moo-chi-she-zha2012} and, more recently, multiple bodies whose shape evolution is coupled through the flow \cite{hewett2017evolution, moore2017riemann}.
From this viewpoint, the example of an eroding porous medium presents new opportunities and challenges. First, since the medium is composed of many individual constituents, multi-body interactions must be considered. Second, the long-range effects inherent to Stokes flow implies that the shape-evolution of bodies are even more strongly coupled than in any other regime.

@@

In closely related work, Mitchell \& Spagnolie used a traction integral equation formulation to simulate viscous erosion of a single body in three dimensions \cite{mit-spa2016}. Their formulation was flexible enough to simulate a {\em mobile} body that erodes in various background flows (uniform flow, shear flow, etc.). An advantage of the traction formulation is that is directly provides the shear stress that is needed to evolve surfaces. 
In the present work, we aim to simulate the erosion of a porous medium and further we choose to fix the bodies to their initial locations. Thus, instead than simulating a single, mobile body as in \cite{mit-spa2016}, we will simulate the erosion of many, immobile bodies. Furthermore, we will employ a standard double-layer formulation, as it allows us to easily calculate bulk quantities such as vorticity, pressure, and velocity in a post-processing step.
